---
title: "R Notebook"
output: html_notebook
authors: "Andrew Rost & Pavel Sachivichik"
purpose: "Financial Analysis of Oil Companies"
date: "12/5/2019"
---
Topic:
Our topic for this final project is the financial analysis of the oil industries stocks. We decided to do this topic because we think that it is a a good challenge and we will be able to implement the techniques we learned in this class. Both of us follow the stock market and we thought that doing an analysis of the oil industry would be very interesting. The stocks prices within the oil industry is very hecktic and uncertain. We wanted to see if we could find out what kind factors influence the rise and fall of the market. There are many aspects within the oil industry that influence the stock price and we would like to look at many of them. 

Research Questions:
- Have the performance of Oil & Gas stocks been performing better than the market?
- Why are some oil & gas companies doing better than others?
- What influences oil companies to do good or bad in the stock market?
- How does the the stock of service companies correspond compared to how well or how bad oil companies do in the       market?
- How does the price of oil compared to the price of the stock?
- How do accidents affect the stock of oil companies.



Here, we establish which packages we are importing:
```{r message=FALSE, warning=FALSE}
library(readr)
library(DataComputing)
library(dplyr)
library(tidyverse)
library(lubridate)
library(readr)
library(ggplot2)
```

This block is used to import the various tables that are used in our report:
```{r}
CVX <- read_csv("CVX.csv")
head(CVX)

HAL <- read_csv("HAL.csv")
head(HAL)

XOM <- read_csv("XOM.csv")
head(XOM)

SP <- read_csv("GSPC.csv")

BrentOilPrices <- read_csv("BrentOilPrices.csv")


BrentOilPrices <-
  BrentOilPrices %>%
  mutate(Date = mdy(Date))
```

This graph describes how a few oil/gas companies (CVX - Chevron, HAL - Haliburton) are performing when compared to the S&P 500 (SP):
```{r}
adjustS <- 2790.37/120.67
adjustH <- 32.66/120.67
combined <-
CVX %>%
inner_join(HAL, by=c("Date"="Date")) %>%
inner_join(SP, by=c("Date"="Date")) %>%
filter(Date>="2018-12-01") %>%
mutate(CVX = Close.x) %>%
mutate(SP= Close/adjustS) %>%
mutate(HAL = Close.y/adjustH) %>%
select(Date, CVX, SP, HAL)
tail(combined)
g <- ggplot(data = combined) + geom_line(aes(x = Date, y=CVX, color = "CVX")) +
geom_line( aes(x=Date, y = combined$SP,color = "SP")) +
geom_line(aes(x=Date, y = combined$HAL,color = "HAL")) + ylab("Closing Price") + scale_colour_manual(name='', values=c('SP'= "red",
'HAL'='blue', "CVX"="black"))
g
```

This graph is used to compare the performance of HAL stock against a basic oil index
```{r}
adjustO <- 60.17/32.66
combined2 <-
HAL %>%
inner_join(BrentOilPrices, by=c("Date"="Date")) %>%
filter(Date>="2018-12-01") %>%
mutate(HAL = Close) %>%
mutate(Oil= Price/adjustO) %>%
select(Date, HAL, Oil)

g <- ggplot(data = combined2) + geom_line(aes(x = Date, y = HAL, color = "HAL")) +
geom_line(aes(x=Date, y = Oil, color = "Oil")) + ylab("Price") + scale_colour_manual(name='', values=c('Oil'= "red", "HAL"="black"))
g
```

This graph is used to compare the performance of CVX stock against a basic oil index
```{r}
adjustO <- 60.17/120.67
combined3 <-
CVX %>%
inner_join(BrentOilPrices, by=c("Date"="Date")) %>%
filter(Date>="2018-12-01") %>%
mutate(CVX = Close) %>%
mutate(Oil= Price/adjustO) %>%
select(Date, CVX, Oil)
g <- ggplot(data = combined3) + geom_line(aes(x = Date, y = CVX, color = "CVX")) +
geom_line(aes(x=Date, y = Oil, color = "Oil")) + ylab("Price") + scale_colour_manual(name='', values=c('Oil'= "red", "CVX"="black"))
g
```

This graph is used to compare the performance of XOM stock against a basic oil index
```{r}
adjustO <- 60.17/81.22
combined4 <-
XOM %>%
inner_join(BrentOilPrices, by=c("Date"="Date")) %>%
filter(Date>="2018-12-01") %>%
mutate(XOM = Close) %>%
mutate(Oil= Price/adjustO) %>%
select(Date, XOM, Oil)
g <- ggplot(data = combined4) + geom_line(aes(x = Date, y = XOM, color = "XOM")) +
geom_line(aes(x=Date, y = Oil, color = "Oil")) + ylab("Price") + scale_colour_manual(name='', values=c('Oil'= "red", "XOM"="black"))
g
```


